Design Document for Dimension Reduction in R Package glmm

### 2020-07-16 DesignDocs/DesignDocDimReduction.tex
tweaks to dim reduction design doc

I am figuring out one of these two options:
\begin{itemize}
\item[Plan A:] Change PQL so that similar random effects are collapsed into a shared, single random effect.
\item[Plan B:] Keep the PQL optimization as it is, and add another step in between the two steps. At this time, the similar random effects are collapsed into a shared, single random effect.
  \end{itemize}

  I'm not sure, but it seems like Plan A is better than Plan B. It makes sense that we'd want to try collapsing the random effects while making sure this simplification not ruin the fit of the model.

   In the currently used version of PQL, there is an optimization within an optimization. The inner optimization estimates the fixed effects and predicts the random effects. The outer optimization works on the variance components. \textbf{I believe that adding a penalty term to the inner loop will collapse the random effects.} We want the random effects to be similar to one other, so the distance (or the square of the distance, if we prefer a ridge-like penalty) between the random effects must be controlled.

@@

There are two variations of PQL, both of which are described in the vignette
 \texttt{re.pdf} in the R package \texttt{aster} (Geyer, 2014). Both variations have an inner optimization and an outer optimization. I use the variation of PQL that is not the original version, but is pretty close and is better behaved.  In this version, the inner optimization finds $\tilde{\beta}$ and $\tilde{s}$ given $X$, $Z$ and $A$. Then, given $\tilde{\beta}$ and $\tilde{s}$, the outer optimization finds $A$. \textbf{I believe that adding a penalty term to the inner loop will collapse the random effects. We want the random effects to be similar to one other, so the distance (or the square of the distance) between the random effects must be controlled. In other words, if $s$ has components $s_1, \ldots , s_q$, then we want to have penalty term $\sum_{k \neq j} |s_k - s_j|$ if we want a lasso-like penalty or $\sum_{k \neq j} (s_k - s_j)^2$ if we want a ridge-like penalty.}

@@

The inner optimization maximizes the penalized log likelihood. Defining
\begin{align}
\eta=X\beta +ZAs
\end{align}
we calculate the  log likelihood as 
\begin{align}
l(\eta)= y' \eta - c(\eta) 
\end{align}
and the penalized likelihood (pre-collapsing) as:
\begin{align}
 l(\eta)- \dfrac{1}{2} s's.
\end{align}
Then our doubly-penalized likelihood is
\begin{align}
 l(\eta)- \dfrac{1}{2} s's. - \lambda \sum_{k \neq j} (s_k - s_j)^2
\end{align}
where $\lambda$ is some number we can tweak to force more/less collapsing.

I HAVE NOT REDONE THESE CALCULATIONS BELOw

>>>>

I am figuring out one of these two options:
\begin{itemize}
\item[Plan A:] Change PQL so that similar random effects are fused into a shared, single random effect.
\item[Plan B:] Keep the PQL optimization as it is, and add another step in between the two steps. At this time, the similar random effects are fused into a shared, single random effect.
  \end{itemize}

  I'm not sure, but it seems like Plan A is better than Plan B. It makes sense that we'd want to try fusing the random effects while making sure this simplification not ruin the fit of the model.

   In the currently used version of PQL, there is an optimization within an optimization. The inner optimization estimates the fixed effects and predicts the random effects. The outer optimization works on the variance components. \textbf{I believe that adding a penalty term to the inner loop will fuse the random effects.} We want the random effects to be similar to one other, so the distance (or the square of the distance, if we prefer a ridge-like penalty) between the random effects must be controlled. This is almost exactly pairwise fusion lasso (Petry et. al. 2011), except we are fusing \textit{random} effects instead of \textit{fixed} effects.

@@

There are two variations of PQL, both of which are described in the vignette
 \texttt{re.pdf} in the R package \texttt{aster} (Geyer, 2014). Both variations have an inner optimization and an outer optimization. I use the variation of PQL that is not the original version, but is pretty close and is better behaved.  In this version, the inner optimization finds $\tilde{\beta}$ and $\tilde{s}$ given $X$, $Z$ and $A$. Then, given $\tilde{\beta}$ and $\tilde{s}$, the outer optimization finds $A$. \textbf{I believe that adding a penalty term to the inner loop will fuse the random effects. We want the random effects to be similar to one other, so the distance (or the square of the distance) between the random effects must be controlled. In other words, if $s$ has components $s_1, \ldots , s_q$, then we want to have penalty term $\sum_{k \neq j} |s_k - s_j|$ if we want a lasso-like penalty or $\sum_{k \neq j} (s_k - s_j)^2$ if we want a ridge-like penalty.}

@@

The inner optimization maximizes the penalized log likelihood. Defining
\begin{align}
\eta=X\beta +ZAs
\end{align}
we calculate the  log likelihood as 
\begin{align}
l(\eta)= y' \eta - c(\eta) 
\end{align}
and the penalized likelihood (pre-fusion) as:
\begin{align}
 l(\eta)- \dfrac{1}{2} s's.
\end{align}
Then our doubly-penalized likelihood is
\begin{align}
 l(\eta)- \dfrac{1}{2} s's. - \lambda \sum_{k \neq j} (s_k - s_j)^2
\end{align}
where $\lambda$ is some number we can tweak to force more/less fusion.

I HAVE NOT REDONE THESE CALCULATIONS BELOW and it will kind of suck because of the new fusion penalty.
